\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-20 paper.qmd
Add hexbin notation vis

The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $ C_{h}^{(2)} = (c_{h1}, c_{h2})$. The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid. The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

@@

```{r}
hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = r2), fill = "white", color = "black", alpha = 0, linewidth = 1) +
    geom_segment(x = 0, y = -0.35, xend = 1, yend = -0.35,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = r2,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
  coord_fixed()
```

@@

Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, $b = b_1 \times b_2$. To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$. The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $n$. The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-mse-scurve-b).

@@

<!--To generate errors for different total number of bins-->
```{r}
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 2:27 #sqrt(NROW(training_data_scurve))

@@

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 50%

>>>>

The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$. The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid. The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

@@

```{r}
hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2- 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
  coord_fixed()
```

@@

To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$. The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$. The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-mse-scurve-b).

@@

<!--To generate errors for different total number of bins-->
```{r}
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)

@@

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 20%

### 2024-05-20 paper.tex
Add hexbin notation vis

The 2D hexagon grid is defined by the number of bins in each direction
\((b_1, b_2)\), giving total number of bins as \(b = b_1 \times b_2\),
and hexagon id, \(h = 1, \dots, b\). Each hexagon, \(H_h\) is uniquely
described by centroid, \$ C\_\{h\}\^{}\{(2)\} = (c\_\{h1\},
c\_\{h2\})\$. The lower left position where the grid starts at
\((s_1, s_2)\), which correspond to the lowest left centroid. The values
of \(s_i\) need to be below their respective minimum variable values,
and could be a full bin lower, to allow a buffer (\(q\)) corresponding
to a full hexagon width (\(a_1\)) and height (\(a_2\)) around the data.
The values of \(b_i\) are variables to be computed that define the
reduction in size of the data (\(n\) to \(m\)).

@@

Furthermore, the total number of bins is calculated based on the number
of bins along the x-axis and y-axis, \(b = b_1 \times b_2\). To
determine the effective \(b\), candidate values are selected based on
the range between the minimum and approximate maximum \(b_1\), because
\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
while the maximum number is estimated by taking the square root of
\(n\). The analysis evaluates the MSE across varying \(b\) within this
range, covering the minimum to maximum values along both axes (see
Figure~\ref{fig-mse-scurve-b}).

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}

>>>>

The \gD{} hexagon grid is defined by the number of bins in each
direction \((b_1, b_2)\), giving total number of bins as
\(b = b_1 \times b_2\), and hexagon id, \(h = 1, \dots, b\). Each
hexagon, \(H_h\) is uniquely described by centroid,
\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). The lower left position where the
grid starts at \((s_1, s_2)\), which correspond to the lowest left
centroid. The values of \(s_i\) need to be below their respective
minimum variable values, and could be a full bin lower, to allow a
buffer (\(q\)) corresponding to a full hexagon width (\(a_1\)) and
height (\(a_2\)) around the data. The values of \(b_i\) are variables to
be computed that define the reduction in size of the data (\(n\) to
\(m\)).

@@

To determine the effective \(b\), candidate values are selected based on
the range between the minimum and approximate maximum \(b_1\), because
\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
while the maximum number is estimated by taking the square root of
\(\frac{n}{2}\). The analysis evaluates the MSE across varying \(b\)
within this range, covering the minimum to maximum values along both
axes (see Figure~\ref{fig-mse-scurve-b}).

@@

\includegraphics[width=1\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
